Model Finding for Recursive Functions in SMT\thanks{ This research is partially supported by the Inria technological development action ``Contre-exemples Utilisables par Isabelle et Coq'' (CUIC). }

### 2015-05-07 tex/main.tex
more

\begin{abstract}
SMT solvers have recently been extended with techniques for finding models
in presence of (universally) quantified formulas.
However, these generally cannot cope with universal quantifiers ranging over
infinite types, such as integers and algebraic datatypes. This paper introduces
a translation that eliminates infinite quantifiers in axioms specifying a large
class of recursive functions, including terminating, tail-recursive, and
productive functions. An empirical evaluation confirms that the approach works
well on benchmarks from two sources. The translation is implemented as a
preprocessor in the solver CVC4.
\end{abstract}

@@

In this paper, we present a method for translating formulas involving recursice function
definitions into formulas where finite model finding techniques can be applied.
The recursive functions must meet a semantic criterion to be admissible
(Section~\ref{sec:prelim}). This criterion is met by terminating recursive
function definitions, but also by tail-recursive functions and even by
\relax{productive} corecursive functions. It is, however, not met by inconsistent
definitions such as $\forall n : {\typeint}.\;\, \con{f}(n) \teq \con{f}(n) +
1$, %ct and it is possible to construct consistent examples that do not meet it.
as well as other example of consistent definition.

@@

>>>>

\begin{abstract}
SMT solvers have recently been extended with techniques for finding models
in presence of (universally) quantified formulas.
However, these generally cannot cope with universal quantifiers ranging over
infinite types, such as integers and algebraic datatypes. This paper introduces
a translation that eliminates infinite quantifiers in axioms specifying a large
class of recursive functions, including terminating and
productive functions. An empirical evaluation confirms that the approach works
well on benchmarks from two sources. The translation is implemented as a
preprocessor in the solver CVC4.
\end{abstract}

@@

In this paper, we present a method for translating formulas involving recursice function
definitions into formulas where finite model finding techniques can be applied.
The recursive functions must meet a semantic criterion to be admissible
(Section~\ref{sec:prelim}). This criterion is met by terminating recursive
function definitions, but also by
\relax{productive} corecursive functions. It is, however, not met by inconsistent
definitions such as $\forall n : {\typeint}.\;\, \con{f}(n) \teq \con{f}(n) +
1$, %ct and it is possible to construct consistent examples that do not meet it.
as well as other example of consistent definition.

@@

Admissibility is a semantic criterion that must be satisfied for each function
definition before applying the translation described in
Section~\ref{sec:encoding}. This criterion is general enough to accommodate
\emph{well-founded} function definitions, also called ``terminating'' definitions in a
slight abuse of terminology. In such definitions, all recursive calls are
decreasing with respect to a well-founded relation $r$, which must be supplied
by the user or inferred automatically using a termination prover.

First-order logic has built-in notion of computation or termination. To ensure
that a function specification is well-founded, it is sufficient to require that
it would terminate when seen as a functional program for \relax{some}
evaluation order. For example,
$\{\forall n : {\typeint}.\;\, \lite\bigl(
n \leq 0,\allowbreak\; \con{p}( n ) \teq 1,\allowbreak\; \con{p}( n ) \teq 2 * \con{p}( n - 1 ) \bigr)\}$
can be shown well-founded under an evaluation strategy that consides one of the
two $\lite$ branches, based on the result of evaluating the condition.
Logically, such dependencies can be captured by congruence rules. Krauss
develops these ideas systematically in the more general context of higher-order
logic \cite[Section 2]{krauss-2009-phd}.

\begin{theorem}
Well-founded f
\end{theorem}

\newpage
